---
title: "Exercise 1"
output: pdf_document
---


This first example gives you a short introduction to the back-door criterion and how to use it in R
We begin by loading necessary packages - if you don't have them, install them now.


```{r, message=FALSE, warning=FALSE}
library(ggdag)
library(dagitty)
```

We will define a DAG using the dagitty command. The reason why we need a long statement in quotation marks is that 
the dagitty command actually takes a dag statement from the dagitty website (http://www.dagitty.net).
For this example, I created the graph in dagitty, and then I just copied and pasted the model code into Rstudio.
I did not type out this whole graph below. If you had to, it works as follows:
- every variable is on a single line - the pos argument is optional, but allows you to control the position of each node
- arrows between variables are simply done by typing two variables and putting an arrow between them

```{r}
dag1 <- dagitty('dag {
             Coach [pos="-4.392,-7.906"]
             ConnectiveTissueDisorder [pos="3.494,-5.099"]
             ContactSport [pos="-1.157,2.548"]
             FitnessLevel [pos="-1.489,-4.530"]
             Genetics [pos="2.022,-7.906"]
             Injury [outcome,pos="4.969,8.605"]
             IntraGameProprioception [pos="-1.196,8.487"]
             NeuromuscularFatigue [pos="1.501,-1.235"]
             PreGameProprioception [pos="-3.638,-1.235"]
             PreviousInjury [pos="-4.293,4.175"]
             TeamMotivation [pos="-7.175,-0.950"]
             TissueWeakness [pos="4.236,1.857"]
             WarmUpExercises [exposure,pos="-7.000,8.650"]
             Coach -> FitnessLevel
             Coach -> TeamMotivation
             ConnectiveTissueDisorder -> NeuromuscularFatigue
             ConnectiveTissueDisorder -> TissueWeakness
             ContactSport -> IntraGameProprioception
             ContactSport -> PreviousInjury
             FitnessLevel -> NeuromuscularFatigue
             FitnessLevel -> PreGameProprioception
             Genetics -> ConnectiveTissueDisorder
             Genetics -> FitnessLevel
             Genetics -> NeuromuscularFatigue
             IntraGameProprioception -> Injury
             NeuromuscularFatigue -> Injury
             NeuromuscularFatigue -> IntraGameProprioception
             PreGameProprioception -> WarmUpExercises
             TeamMotivation -> PreviousInjury
             TeamMotivation -> WarmUpExercises
             TissueWeakness -> Injury
             WarmUpExercises -> IntraGameProprioception
             }')
```


The ggdag package allows plotting of the model, and also of adjustment sets that are identified using the back-door criterion. Because the model is large, and there are many sets, the adjustment set display is not very helpful in this case. 


```{r}
ggdag(dag1,text_size = 3,text_col = "orange")
ggdag_adjustment_set(dag1, text_size = 3, text_col = "black")
```


The adjustmentSets command, simply lists the adjustment sets. 

```{r}
adjustmentSets(dag1)
```



Exercise: 

1.) Using the daggity language, create the following DAG in R.

2.) Using the ggdag package, plot the DAG that you just created. 



```{r, echo=FALSE}
dag2 <- dagitty('dag {
A [pos="-2.200,-1.520"]
B [pos="1.400,-1.460"]
D [outcome,pos="1.400,1.621"]
E [exposure,pos="-2.200,1.597"]
Z [pos="-0.300,-0.082"]
A -> E
A -> Z
B -> D
B -> Z
E -> D
Z -> D
Z -> E
}
')

ggdag(dag2,text_size = 3,text_col = "orange")
```


3.) Use the back-door criterion by hand to find an adjustment set for the causal effect of E on D. 

4.) Confirm this set by using the adjustmentSets command. 


